The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2024-08-12 ms/ms.tex
overleaf

\section{Introduction}
Water availability, land temperature extremes, fire risk, and crop productivity will all see impacts from climate change, and are among the many processes represented within the terrestrial components of Earth System Models. Understanding how these processes respond to and influence CO$_2$ concentrations is a critical facet of climate change research. Uncertainty in climate model projections varies by domain and generally increases with extended time horizons \cite{koven2022}. Centennial-scale estimates of the cumulative terrestrial carbon sink have been especially challenging, with high uncertainty persisting across model generations \cite{friedlingstein2014,arora2020}. Something about water \cite{zarakas2024}. A portion of this uncertainty is irreducible, inherent to the challenge of predicting vegetation dynamics in a novel climate \cite{lovenduski2017}. Still, our expectation is that some of this uncertainty is indeed reducible, if we can more effectively utilize the ongoing expansion of observational data sources from remote sensing, meteorological stations, flux towers, and field campaigns. The capability to efficiently ingest these data and improve simulation performance is especially valuable for actionable science \cite{cheng2023}. However, given increasingly comprehensive land modeling systems, and the wide array of observational products, several technical hurdles exist that hinder effective model development and calibration.

@@

In this project we create a large PPE with the Community Terrestrial Systems Model (CTSM) \cite{lawrence2019}, in the Community Land Model with biogeochemistry configuration (CLM-BGC). We extend the work of \citeA{dagon2020} which utilized a PPE to calibrate a subset of CLM parameters. Here we include a broader suite of parameters and continue to enhance the software tools for parameter perturbation and optimization.
The goal of this project is to systematize the parameter perturbation process within the CTSM modeling framework, and develop the necessary tools and datasets to efficiently test parameter effects across a wide range of land model processes. 
In doing so, we have generated a large PPE, comprising thousands of parameter sensitivity tests.
In this paper we present that dataset, describe how it was produced, and survey some potential applications.
The dataset has already demonstrated utility for diagnosing parameter effects \cite{cheng2023,yan2023a,yan2023b}, while the software and modeling infrastructure has greatly expanded our capability to generate insights about parameter effects and uncertainties within CTSM/CLM.

>>>>

\section{Introduction}
Water availability, land temperature extremes, fire risk, and crop productivity will all see impacts from climate change, and are among the many processes represented within the terrestrial components of Earth System Models. Understanding how these processes respond to and influence CO$_2$ concentrations is a critical facet of climate change research. Uncertainty in climate model projections varies by domain and generally increases with extended time horizons \cite{koven2022}. Land processes substantially influence climate directly through, for example, evapotranspiration \cite{zarakas2024}, and indirectly through carbon-climate feedbacks.
Making centennial-scale projections of the cumulative terrestrial carbon sink gas been especially challenging, with high uncertainty persisting across model generations \cite{friedlingstein2014,arora2020}, and limited efficacy of emergent constraints. A portion of this uncertainty is irreducible, inherent to the challenge of predicting vegetation dynamics in a novel climate \cite{lovenduski2017}. Still, our expectation is that some of this uncertainty is indeed reducible, if we can more effectively utilize the ongoing expansion of observational data sources from remote sensing, meteorological stations, flux towers, and field campaigns. The capability to efficiently ingest these data and improve simulation performance is especially valuable for actionable science \cite{cheng2023}. However, given increasingly comprehensive land modeling systems, and the wide array of observational products, several technical hurdles exist that hinder effective model development and calibration.

@@

In this project we create a large PPE with the Community Terrestrial Systems Model (CTSM) \cite{lawrence2019}, in the Community Land Model with biogeochemistry configuration (CLM-BGC). We extend the work of \citeA{dagon2020} which utilized a PPE to calibrate a subset of CLM parameters. Here we include a broader suite of parameters and continue to enhance the software tools for parameter perturbation and optimization.
The goal of this project is to systematize the parameter perturbation process within the CTSM modeling framework, and develop the necessary tools and datasets to efficiently test parameter effects across a wide range of land model processes. 
In doing so, we have generated a large PPE, comprising thousands of parameter sensitivity tests.
In this paper we present that dataset, describe how it was produced, and survey some potential applications.
The dataset has already demonstrated utility for diagnosing parameter effects \cite{cheng2023,yan2023a,yan2023b,zarakas2024}, while the software and modeling infrastructure has greatly expanded our capability to generate insights about parameter effects and uncertainties within CTSM/CLM.
